---
title: "Visualization and EDA"
author: "Grace Liu"
date: "10/11/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

devtools::install_github("p8105/p8105.datasets") # install.packages("devtools")
library(p8105.datasets)
```

## Problem 1
```{r}
brfss_data = brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(topic == "Overall Health")

brfss_data$response = factor(c("Excellent", "Very good", "Good", "Fair", "Poor"))
```

**In 2002, which states were observed at 7 locations? **
```{r}
brfss_data %>% 
  group_by(year, locationabbr) %>% 
  filter(year == 2002) %>% 
  summarize(n_locations = n_distinct(locationdesc) == 7) %>% 
  filter(n_locations == TRUE)
```
In 2002, there were 3 states where 7 locations were observed.  

**“Spaghetti plot” showing the number of locations in each state from 2002 to 2010**
```{r}
brfss_data %>%
  group_by(year, locationabbr) %>%
  count(n_locations = n_distinct(locationdesc)) %>% 
  ggplot(aes(x = year, y = n_locations, color = locationabbr)) + 
    geom_smooth(se = FALSE, alpha = .5) +
    labs(
      title = "Number of locations in each state, 2002-2010",
      x = "Year",
      y = "Number of locations") + 
    theme_minimal()
```

Make a table showing, for the years 2002, 2006, and 2010, the mean and standard deviation of the proportion of “Excellent” responses across locations in NY State.


For each year and state, compute the average proportion in each response category (taking the average across locations in a state). Make a five-panel plot that shows, for each response category separately, the distribution of these state-level averages over time.

